Supporting online material for\\ \emph{Effectiveness of UNAIDS targets and HIV vaccination across 127 countries}

### 2016-08-20 Documents/supplementary_text.tex
Hacking on supplement.

\documentclass[11pt]{article}

@@

We developed a continuous-time, compartmental model that divides
people aged $15$\;y and older into non-overlapping HIV states.  We
chose to consider only these ages because they are the vast majority
of PLHIV, 93\% in 2014 \cite{UNICEF}, and because HIV and demographic
data are often stratified into these ages, although some data used to
parametrize the model was only available for ages $15$--$49$ rather
than for ages $15$ and older.  The background mortality rate ($\mu$)
in each of the 127 countries was collected from published estimates
\cite{World_Development_Indicators2013-ee}.  The recruitment rate into
ages $15\;\text{y}$ and older group ($B$) was set by the difference
between the country's population growth rate \cite{WorldBankpg} and
its background mortality rate
\cite{World_Development_Indicators2013-ee}, as would be true if the
population was at its stable age structure.

@@

\begin{table}
  \begin{center}
    \begin{tabularx}{\textwidth}{lXlll}
      \hline
      & Definition & Value & Reference \\
      \hline
      $\mu$ & Background death rate & Country specific, see Table S2
      & \cite{World_Development_Indicators2013-ee} \\
      $B$ & Recruitment rate & Country specific, see Table S2
      & \cite{World_Development_Indicators2013-ee, WorldBankpg} \\
      $\delta$	& Rate of leaving acute infection
      & $\Triangular$(2, 4.14, 9.6)\;y$^{-1}$
      & \cite{Hollingsworth2008-iy} \\
      $\sigma$	& Rate of developing AIDS without viral suppression
      & 0.1064\;y$^{-1}$ & \cite{Morgan2002-cq} \\
      $\theta$ & Rate of developing AIDS with viral suppression
      & See eq.~(\ref{theta}) & --- \\
      $\gamma$ & Rate of viral suppression
      & $\Uniform$(0.5, 1.5)\;y$^{-1}$
      & \cite{Currie2009-yz} \\
      $\nu$ & Death rate from AIDS & 0.5\;y$^{-1}$
      & \cite{Morgan2002-cq} \\
      $\rho$ & Vaccine efficacy & 50\%\,[30\%, 70\%] & --- \\
      $\omega$	& Reduction in lifetime with viral suppression
      & $\Uniform$(5, 8)\;y
      & \cite{Unaids2014-ue, Samji2013-kf} \\
      $\tau_{A}$ & Transmissibility during acute phase
      & $\Triangular$(0.0039, 0.0082, 0.0150)
      & \cite{Skarbinski2015-ni,Wawer2005-us}\\
      $\tau_{U}$ & Transmissibility after acute phase
      & $\Triangular$(0.00077, 0.0014, 0.00251)
      & \cite{Hughes2012-so} \\
      $\varepsilon$ & Relative transmissibility with
      viral suppression & $\BetaPERT$(0.08, 0.002, 0.57)
      & \cite{Donnell2010-xo} \\
      $n$ & Coital acts per year & $\Uniform$(96, 108)
      & \cite{Wawer2005-us,Abdool_Karim2010-cm}\\
      $\alpha$ & Diagnosis rate & See eq.~(\ref{diagnosis_rate}) & --- \\
      $\phi$ & Treatment rate & See eq.~(\ref{treatment_rate}) & --- \\
      $\psi$ & Rate of relapse to untreated & See eq.~(\ref{relapse_rate})
      & --- \\
      $\zeta$ & Vaccination rate & See eqs.~(\ref{vaccination_rate}) & --- \\
      \hline
    \end{tabularx}
    \caption{Model parameters. $\Triangular$, $\Uniform$, and
      $\BetaPERT$ are sampling distributions: see
      \autoref{uncertainty} for a full description.  For vaccine
      efficacy, $50\%$ was the baseline, and $30\%$ and $70\%$ were
      also used in scenario analysis.}
    \label{model_param}
  \end{center}
\end{table}

People susceptible to HIV ($S$) progress to acute infection ($A$)
according to the force of infection ($\lambda$,
eq.~\eqref{force_of_infection}), which depends on the HIV transmission
rate, estimated from the country-specific incidence and prevalence
data (\autoref{model_fitting}), adjusted for the current number of
acutely infected, untreated or ineffectively treated, and viral
suppression PLHIV.  Acute infection is characterized by high viral
loads and lasts an average of 2.90 [95\% confidence interval 1.23,
6.00]\;months \cite{Hollingsworth2008-iy}.  (We converted these
durations to rates like $1 / 2.90\;\text{month$^{-1}$} \approx
4.14\;\text{y$^{-1}$}$ and set the parameter using random samples from
$\Triangular(2, 4.14, 9.6)\;\text{y$^{-1}$}$ to capture the
uncertainty in its value.  See \autoref{uncertainty} for the
definition of $\Triangular(a, b, c)$ and other distributions.)  We
assumed that people remain undiagnosed during their acute infection
phase. After acute infection, people move into the chronic undiagnosed
class ($U$).  People in the undiagnosed class will be diagnosed at the
rate $\alpha$, which is determined by the country's current diagnosis
level and its diagnosis target (\autoref{targets}).  Diagnosed people
($D$) transition to the treated compartment ($T$) upon starting ART at
the rate $\phi$, dependent on the country's current treatment level
and its treatment target (\autoref{targets}).  After some time,
treated people then transition to the viral-suppression class ($V$),
at the rate $\gamma$.  We took $\gamma$ as a random sample from
$\Uniform(0.5, 1.5)\;\text{y$^{-1}$}$, that is, viral suppression
occurs after between 8 months and 2 years on treatment
\cite{Currie2009-yz}.  The disengagement from treatment moves people
from treated ($T$) and viral suppression ($V$) at the rate $\psi$,
dependent on the current level of viral suppression and the country's
target (\autoref{targets}), back to the diagnosed class ($D$).

@@

In scenarios with vaccine, susceptible people are vaccinated at rate
$\zeta$, depending on the country's current vaccination coverage and
its coverage target (\autoref{targets}), transitioning into the
vaccinated compartment ($R$), where the vaccine reduces their force of
infection by the efficacy factor $\rho$.  Without vaccine,
$\zeta = 0$.

The model equations are
\begin{equation}
  \label{model_eqns}
  \begin{split}
    \frac{\md S}{\md t} &= B N - \lambda S - \zeta S- \mu S,
    \\
    \frac{\md R}{\md t} & = \zeta S - (1 - \rho) \lambda R - \mu R,
    \\
    \frac{\md A}{\md t} &= \lambda S + (1 - \rho) \lambda R - \delta A - \mu A,
    \\
    \frac{\md U}{\md t} &= \delta A - \alpha U - \mu U - \sigma U,
    \\
    \frac{\md D}{\md t} &=  \alpha U + \psi T + \psi V
    - \phi D - \mu D - \sigma D,
    \\
    \frac{\md T}{\md t} &= \phi D - \psi T - \gamma T - \mu T
    - \sigma T,
    \\
    \frac{\md V}{\md t} &= \gamma T - \psi V - \mu V - \theta V,
    \\
    \frac{\md W}{\md t} &= \sigma U + \sigma D + \sigma T + \theta V -
    \nu W,
  \end{split}
\end{equation}
with force of infection
\begin{equation}
  \label{force_of_infection}
  \begin{split}
    \lambda &= \frac{\beta_A A + \beta_U (U + D + T) + \beta_T V}{N},
    \\
    N &= S + R +  A + U + D + T + V,
    \\
    \beta_x &= \eta \left[1 - (1 - \tau_x)^n\right]
    \text{ for $x \in \{A, U, T\}$},
    \\
    \tau_T &= \varepsilon \tau_U,
  \end{split}
\end{equation}
where $\eta$ is the transmission rate (\autoref{model_fitting}).

\comment{Describe transmissibilities.}

\comment{Initial conditions.}
The number of people aged 15--49 years \cite{The_World_Bank2016-fd}
and the proportion diagnosed, treated, and with viral suppression
(\autoref{data_sources}) were used to initialize the model for year
2015.  Due to a lack of data on acute infections, and since the acute
phase is so much shorter than the other stages, we assumed that
initially there were no acute infections.

@@

The current proportion of PLHIV who know their diagnosis is
\begin{equation}
  p_D(t) = \frac{D(t) + T(t) + V(t) + W(t)}
  {A(t) + U(t) + D(t) + T(t) + V(t) + W(t)},
\end{equation}
and the desired target for the current level of diagnosis is
$p_D^*(t)$.  For the rate of new diagnosis, we used the function
\begin{equation}
  \label{diagnosis_rate}
  \alpha(t) = \alpha_{\max} H\!\left(p_D^*(t) - p_D(t)\right),
\end{equation}
where $H(x)$ is the Heaviside-like function
\begin{equation}
  H(x) =
  \begin{cases}
    0 & \text{if $x < 0$},
    \\
    x / \chi & \text{if $0 \leq x \leq \chi$},
    \\
    1 & \text{if $x > \chi$},
  \end{cases}
\end{equation}
with $\chi = 0.001$.  The linear segment connecting $H = 0$ and
$H = 1$ was used to avoid difficulties in computing numerical
solutions that occur when using a discontinuous function.  This
function allows new diagnoses ($\alpha(t) > 0$) when the current
diagnosis level is below the target (i.e., when $p_D^*(t) > p_D(t)$),
and stops new diagnoses ($\alpha(t) = 0$) when the current diagnosis
level is at or above target.  We took
$\alpha_{\max} = 1\;\text{y$^{-1}$}$.

The proportion of diagnosed people who are treated is
\begin{equation}
  p_T(t) = \frac{T(t) + V(t) + W(t)}{D(t) + T(t) + V(t) + W(t)}.
\end{equation}
Like the rate of new diagnosis, the rate of enrolling new people in
treatment is
\begin{equation}
  \label{treatment_rate}
  \phi(t) = \phi_{\max} H\!\left(p_T^*(t) - p_T(t)\right),
\end{equation}
with $\phi_{\max} = 10$, allowing new treatment ($\phi(t) > 0$) when
the current treatment level is below the target ($p_T^*(t) > p_T(t)$)
and stopping new treatment ($\phi(t) = 0$) when the current treatment
level is at or above target.

The proportion of people on ART who have achieved viral suppression is
\begin{equation}
  p_V(t) = \frac{V(t)}{T(t) + V(t)}.
\end{equation}
The rate of people relapsing to untreated is
\begin{equation}
  \label{relapse_rate}
  \psi(t) = \psi_{\max} H\!\left(p_V(t) - p_V^*(t)\right),
\end{equation}
with $\psi_{\max} = 1$, which, unlike diagnosis and treatment rates,
prevents relapses ($\psi(t) = 0$) when the current level of viral
suppression is below the target ($p_V^*(t) > p_V(t)$) and allows
relapses ($\psi(t) > 0$) when the current level of viral suppression
is at or above target.  We chose to vary the relapse rate to achieve
the desired level of viral suppression because achieving viral
suppression seems to depend on the time on treatment (parametrized by
our rate $\gamma$), while relapse rates are difficult to estimate.

The proportion of susceptible people vaccinated is
\begin{equation}
  p_R(t) = \frac{R(t)}{S(t) + R(t)}.
\end{equation}
Like diagnosis and treatment, the vaccination rate is
\begin{equation}
  \label{vaccination_rate}
  \zeta(t) = \zeta_{\max} H\!\left(p_R^*(t) - p_R(t)\right),
\end{equation}
with $\zeta_{\max} = 1$, allowing vaccination ($\zeta(t) > 0$) when
the current vaccine coverage level is below the target
($p_R^*(t) > p_R(t)$) and stopping vaccination ($\zeta(t) = 0$) when
the current vaccine coverage is at or above target.

The targets for the proportions $p_D$, $p_T$, and $p_V$ in the
achievement of 90--90--90 all operate under the following assumption:
if the initial level $p_x(0)$ is less than $90\%$, the proportion will
increase linearly in the first 5 years and then stay at $90\%$ for the
remaining time; and if the initial level is at $90\%$ or above, the
proportion will remain constant for the duration of the study:
\begin{equation}
  p_x^*(t)
  = p_x(0) + \left[
    \max\left\{0.9, p_x(0)\right\} - p_x(0)
  \right]
  \min\left\{\frac{t}{5}, 1\right\},
\end{equation}
for $x \in \{D, T, V\}$.

We can implement the vaccination compartment in the same way as the
other controls. When we begin vaccination in 2020 (also 2025 in
scenario analysis), the initial vaccination level is 0\%. The coverage
will increase linearly by 25\% per year (also 10\% per year in
scenario analysis). The coverage will continue to increase linearly
until it reaches 70\% coverage (50\% \& 90\% in scenario analysis) and
then stay constant for the remainder of the time period.  95--95--95
goal is implemented subsequently after meeting 90--90--90 by 2020 by
increasing $p_D$, $p_T$, and $p_V$ to 95--95--95 levels by 2030
linearly.

@@

\section{Source code}

The source code for the model and analysis tools is publicly available
at \url{https://github.com/janmedlock/HIV-95-vaccine/}
\cite{medlock2016-git}.

>>>>

\documentclass{article}

@@

We developed a continuous-time, compartmental model that divides
people aged $15$\;y and older into non-overlapping HIV states.  We
chose to consider only these ages because they are the vast majority
of PLHIV, 93\% in 2014 \cite{UNICEF}, and because HIV and demographic
data are often stratified into these ages, although some data used to
parametrize the model was only available for ages $15$--$49$ rather
than for ages $15$ and older.  The background mortality rate ($\mu$)
in each of the 127 countries was collected from published estimates
\cite{World_Development_Indicators2013-ee}.  The recruitment rate into
ages $15\;\text{y}$ and older group ($\kappa$) was set by the difference
between the country's population growth rate \cite{WorldBankpg} and
its background mortality rate
\cite{World_Development_Indicators2013-ee}, as would be true if the
population was at its stable age structure.

@@

\begin{table}
  \begin{center}
    \begin{tabularx}{\textwidth}{lXlll}
      \hline
      & Definition & Value & Reference \\
      \hline
      $\mu$ & Background death rate & Country specific, Table S2
      & \cite{World_Development_Indicators2013-ee} \\
      $\kappa$ & Recruitment rate & Country specific, Table S2
      & \cite{World_Development_Indicators2013-ee, WorldBankpg} \\
      $\delta$	& Rate of leaving acute infection
      & $\Triangular$(2, 4.14, 9.6)\;y$^{-1}$
      & \cite{Hollingsworth2008-iy} \\
      $\sigma$	& Rate of developing AIDS without viral suppression
      & 0.1064\;y$^{-1}$ & \cite{Morgan2002-cq} \\
      $\theta$ & Rate of developing AIDS with viral suppression
      & Eq.~(\ref{theta}) & --- \\
      $\gamma$ & Rate of viral suppression
      & $\Uniform$(0.5, 1.5)\;y$^{-1}$
      & \cite{Currie2009-yz} \\
      $\nu$ & Death rate from AIDS & 0.5\;y$^{-1}$
      & \cite{Morgan2002-cq} \\
      $\rho$ & Vaccine efficacy & 50\%\,[30\%, 70\%] & --- \\
      $\omega$	& Reduction in lifetime with viral suppression
      & $\Uniform$(5, 8)\;y
      & \cite{Unaids2014-ue, Samji2013-kf} \\
      $\tau_{A}$ & Transmissibility during acute phase
      & $\Triangular$(0.0039, 0.0082, 0.0150)
      & \cite{Skarbinski2015-ni,Wawer2005-us}\\
      $\tau_{U}$ & Transmissibility after acute phase
      & $\Triangular$(0.00077, 0.0014, 0.00251)
      & \cite{Hughes2012-so} \\
      $\varepsilon$ & Relative transmissibility with
      viral suppression & $\BetaPERT$(0.08, 0.002, 0.57)
      & \cite{Donnell2010-xo} \\
      $n$ & Coital acts per year & $\Uniform$(96, 108)
      & \cite{Wawer2005-us,Abdool_Karim2010-cm}\\
      $\alpha$ & Diagnosis rate & See eq.~(\ref{diagnosis_rate}) & --- \\
      $\phi$ & Treatment rate & See eq.~(\ref{treatment_rate}) & --- \\
      $\psi$ & Rate of relapse to untreated & See eq.~(\ref{relapse_rate})
      & --- \\
      $\zeta$ & Vaccination rate & See eqs.~(\ref{vaccination_rate}) & --- \\
      \hline
    \end{tabularx}
    \caption{Model parameters. $\Triangular$, $\Uniform$, and
      $\BetaPERT$ are sampling distributions: see
      \autoref{uncertainty} for a full description.  For vaccine
      efficacy, $50\%$ was the baseline, and $30\%$ and $70\%$ were
      also used in scenario analysis.}
    \label{model_param}
  \end{center}
\end{table}

People susceptible to HIV ($S$) progress to acute infection ($A$)
according to the force of infection ($\lambda$,
eq.~\eqref{force_of_infection}), which depends on the HIV transmission
rate, estimated from the country-specific incidence and prevalence
data (\autoref{model_fitting}), adjusted for the current number of
acutely infected, untreated or ineffectively treated, and viral
suppression PLHIV.  Acute infection is characterized by high viral
loads and lasts an average of 2.90 [95\% confidence interval 1.23,
6.00]\;months \cite{Hollingsworth2008-iy}.  (We converted these
durations to rates like $1 / 2.90\;\text{month$^{-1}$} \approx
4.14\;\text{y$^{-1}$}$ and set the parameter using random samples from
$\Triangular(2, 4.14, 9.6)\;\text{y$^{-1}$}$ to capture the
uncertainty in its value.  See \autoref{uncertainty} for the
definition of $\Triangular(a, b, c)$ and other distributions.)  We
assumed that people remain undiagnosed during their acute infection
phase. After acute infection, people move into the chronic undiagnosed
class ($U$).  People in the undiagnosed class will be diagnosed at the
rate $\alpha$, which is determined by the country's current diagnosis
level and its diagnosis target (\autoref{targets}).  Diagnosed people
($D$) transition to the treated compartment ($T$) upon starting ART at
the rate $\phi$, dependent on the country's current treatment level
and its treatment target (\autoref{targets}).  After some time,
treated people then transition to the viral-suppression class ($V$),
at the rate $\gamma$.  We took $\gamma$ as a random sample from
$\Uniform(0.5, 1.5)\;\text{y$^{-1}$}$, that is, viral suppression
occurs after between 8 months and 2 years on treatment
\cite{Currie2009-yz}.  Disengagement from treatment moves people from
treated ($T$) and viral suppression ($V$) at the rate $\psi$,
dependent on the current level of viral suppression and the country's
target (\autoref{targets}), back to the diagnosed class ($D$).

@@

In scenarios with vaccine, susceptible people are vaccinated at rate
$\zeta$, depending on the country's current vaccination coverage and
its coverage target (\autoref{targets}), transitioning into the
vaccinated compartment ($R$), where the vaccine reduces their force of
infection by the efficacy factor $\rho$.  We took the base-case
efficacy to be $50\%$ and varied it to $30\%$ and $70\%$ in analyzing
vaccination scenarios (\autoref{uncertainty}).  We did not model
specific vaccine dose regime, such that implicit in our efficacy
parameter is the assumption that boosters will be used to maintain
efficacy over time.  Without vaccine, $\zeta = 0$.

We also added differential equations to track the cumulative number of
new infections ($Y$) and AIDS deaths ($Z$) over the model period.

The model equations are
\begin{equation}
  \label{model_eqns}
  \begin{split}
    \frac{\md S}{\md t} &= \kappa N - \lambda S - \zeta S- \mu S,
    \\
    \frac{\md R}{\md t} & = \zeta S - (1 - \rho) \lambda R - \mu R,
    \\
    \frac{\md A}{\md t} &= \lambda S + (1 - \rho) \lambda R - \delta A - \mu A,
    \\
    \frac{\md U}{\md t} &= \delta A - \alpha U - \mu U - \sigma U,
    \\
    \frac{\md D}{\md t} &=  \alpha U + \psi T + \psi V
    - \phi D - \mu D - \sigma D,
    \\
    \frac{\md T}{\md t} &= \phi D - \psi T - \gamma T - \mu T
    - \sigma T,
    \\
    \frac{\md V}{\md t} &= \gamma T - \psi V - \mu V - \theta V,
    \\
    \frac{\md W}{\md t} &= \sigma U + \sigma D + \sigma T + \theta V -
    \nu W,
    \\
    \frac{\md Y}{\md t} &= \lambda \big[S + (1 - \rho) R\big],
    \\
    \frac{\md Z}{\md t} &= \nu W,
  \end{split}
\end{equation}
with force of infection
\begin{equation}
  \label{force_of_infection}
  \begin{split}
    \lambda &= \frac{\beta_A A + \beta_U (U + D + T) + \beta_V V}{N},
    \\
    N &= S + R +  A + U + D + T + V,
    \\
    \beta_A &= \eta \left[1 - (1 - \tau_A)^n\right]
    \\
    \beta_U &= \eta \left[1 - (1 - \tau_U)^n\right]
    \\
    \beta_V &= \eta \left[1 - (1 - \tau_V)^n\right]
    \\
    \tau_V &= \varepsilon \tau_U,
  \end{split}
\end{equation}
where the $\tau$'s are the relative transmissibility of people with
acute ($A$), unsuppressed ($U$ and $T$), and suppressed ($V$)
infections, $n$ is the annual number of coital acts, and $\eta$ is the
country-specific transmission rate (\autoref{model_fitting}).  For
transmissibility during the acute phase, we sampled from
$\Triangular$(0.0039, 0.0082, 0.0150)
\cite{Skarbinski2015-ni,Wawer2005-us}, whereas, for transmissibility
for unsuppressed people, we sampled from
$\Triangular(0.00077, 0.0014, 0.00251)$ \cite{Hughes2012-so}.  The
transmissibility of people with viral suppression relative to those
without suppression has been estimated at 0.08 (95\% confidence
interval 0.002, 0.57) \cite{Donnell2010-xo}: because of the extremely
long left tail, we chose to sample the relative transmissibility
$\varepsilon$ from $\BetaPERT(0.08, 0.002, 0.57)$ rather than a
$\Triangular$ distribution.  (See \autoref{uncertainty} for a full
description of $\BetaPERT(a, b, c)$.)  The annual number of coital $n$
was sampled from $\Uniform(96, 108)$
\cite{Wawer2005-us,Abdool_Karim2010-cm}.

For year 2015, the number of people aged 15--49 years, $M(2015)$; the
prevalence, $p_I(2015)$; the proportion diagnosed, $p_D(2015)$; the
proportion treated, $p_T(2015)$; and proportion with viral
suppression, $p_V(2015)$ (\autoref{data_sources}) were used to
initialize the model.  Due to a lack of data on acute infections, and
since the acute phase is so much shorter than the other stages, we
assumed that initially there were no acute infections.  Due to a lack
of comprehensive data on PLAIDS in each country, we assumed that the
proportion of diagnosed, untreated people ($D + W$) who have AIDS
($W$) in year 2015 was
\begin{equation}
  p_A = \frac{\sigma}{\sigma + \nu},
\end{equation}
which is the equilibrium fraction of diagnosed people ($D + W$) who
have AIDS ($W$) in the absence of treatment and background mortality.
The number initially vaccinated was set to 0.  The initial cumulative
number of new infections and AIDS deaths were set to 0.  Thus, the
model initial conditions are
\begin{equation}
  \label{initial_conditions}
  \begin{split}
    S(2015) &= \big(1 - p_I(2015)\big) M(2015), \\
    R(2015) &= 0, \\
    A(2015) &= 0, \\
    U(2015) &= \big(1 - p_D(2015)\big) p_I(2015) M(2015), \\
    D(2015) &= \big(1 - p_A\big) p_D(2015) \big(1 - p_T(2015)\big)
    p_I(2015) M(2015), \\
    T(2015) &= p_D(2015) p_T(2015) \big(1 - p_V(2015)\big)
    p_I(2015) M(2015), \\
    V(2015) &= p_D(2015) p_T(2015) p_V(2015) p_I(2015) M(2015), \\
    W(2015) &= p_A p_D(2015) \big(1 - p_T(2015)\big) p_I(2015) M(2015), \\
    Y(2015) &= 0, \\
    Z(2015) &= 0.
  \end{split}
\end{equation}

@@

The current proportion of PLHIV who know their diagnosis is
\begin{equation}
  p_D(t) = \frac{D(t) + T(t) + V(t) + W(t)}
  {A(t) + U(t) + D(t) + T(t) + V(t) + W(t)},
\end{equation}
and the desired target for the current level of diagnosis is
$p_D^*(t)$.  For the rate of new diagnosis, we used the function
\begin{equation}
  \label{diagnosis_rate}
  \alpha(t) = \alpha_{\max} H\big(p_D^*(t) - p_D(t)\big),
\end{equation}
where $H(x)$ is the Heaviside-like function
\begin{equation}
  H(x) =
  \begin{cases}
    0 & \text{if $x < 0$},
    \\
    x / \chi & \text{if $0 \leq x \leq \chi$},
    \\
    1 & \text{if $x > \chi$},
  \end{cases}
\end{equation}
with $\chi = 0.001$, which rapidly switches from $H = 0$ when $x < 0$
to $H = 1$ when $x > 0$.  The linear segment connecting $H = 0$ and
$H = 1$ was used to avoid difficulties in computing numerical
solutions that occur when using a discontinuous function.  The
function for the rate of new diagnoses \eqref{diagnosis_rate} allows
new diagnoses ($\alpha(t) > 0$) when the current diagnosis level is
below the target ($p_D^*(t) > p_D(t)$), and stops new
diagnoses ($\alpha(t) = 0$) when the current diagnosis level is at or
above target.  We took $\alpha_{\max} = 1\;\text{y$^{-1}$}$.

The proportion of diagnosed people who are treated is
\begin{equation}
  p_T(t) = \frac{T(t) + V(t) + W(t)}{D(t) + T(t) + V(t) + W(t)}.
\end{equation}
Like the rate of new diagnosis, the rate of enrolling new people in
treatment is
\begin{equation}
  \label{treatment_rate}
  \phi(t) = \phi_{\max} H\big(p_T^*(t) - p_T(t)\big),
\end{equation}
with $\phi_{\max} = 10$, allowing new treatment ($\phi(t) > 0$) when
the current treatment level is below the target ($p_T^*(t) > p_T(t)$)
and stopping new treatment ($\phi(t) = 0$) when the current treatment
level is at or above target.

The proportion of people on ART who have achieved viral suppression is
\begin{equation}
  p_V(t) = \frac{V(t)}{T(t) + V(t)}.
\end{equation}
The rate of people relapsing to untreated is
\begin{equation}
  \label{relapse_rate}
  \psi(t) = \psi_{\max} H\big(p_V(t) - p_V^*(t)\big),
\end{equation}
with $\psi_{\max} = 1$, which, unlike diagnosis and treatment rates,
prevents relapses ($\psi(t) = 0$) when the current level of viral
suppression is below the target ($p_V^*(t) > p_V(t)$) and allows
relapses ($\psi(t) > 0$) when the current level of viral suppression
is at or above target.  We chose to vary the relapse rate to achieve
the desired level of viral suppression because achieving viral
suppression seems to depend on the time on treatment (i.e.,
$1 / \gamma$), while relapse rates are difficult to estimate.

The proportion of susceptible people vaccinated is
\begin{equation}
  p_R(t) = \frac{R(t)}{S(t) + R(t)}.
\end{equation}
Like diagnosis and treatment, the vaccination rate is
\begin{equation}
  \label{vaccination_rate}
  \zeta(t) = \zeta_{\max} H\big(p_R^*(t) - p_R(t)\big),
\end{equation}
with $\zeta_{\max} = 1$, allowing vaccination ($\zeta(t) > 0$) when
the current vaccine coverage level is below the target
($p_R^*(t) > p_R(t)$) and stopping vaccination ($\zeta(t) = 0$) when
the current vaccine coverage is at or above target.

We modeled 3 different targets for the proportion diagnosed ($p_D$), the
proportion treated ($p_T$), and the proportion with viral suppression ($p_V$).
\begin{description}
\item[Status quo] Given initial proportions $p_D(2015)$, $p_T(2015)$,
  and $p_V(2015)$, the targets are for the proportion to remain fixed
  at that level for all times:
  \begin{equation}
    \label{status_quo_target}
    \begin{split}
      p_D^*(t) &= p_D(2015), \\
      p_T^*(t) &= p_T(2015), \\
      p_V^*(t) &= p_V(2015).
    \end{split}
  \end{equation}

\item[UNAIDS 90--90--90] The UNAIDS 90--90--90 target is for $p_D$,
  $p_T$, and $p_V$ to all be raised to $90\%$ by 2020.  We implemented
  this target as linear increases from the 2015 levels, each up to
  $90\%$ in 2020, and constant at $90\%$ from 2020 to 2035, except if
  an initial level is above $90\%$ then it remains constant from 2015
  to 2035 so that meeting the UNAIDS target does not worsen any aspect
  of the treatment cascade.  That is,
  \begin{equation}
    \label{unaids90_targets}
    \begin{split}
      p_D^*(t) &=
      \begin{cases}
        F\big(t, 2015, p_D(2015), 2020, 0.9\big)
        & \text{if $p_D(2015) < 0.9$},
        \\
        p_D(2015) & \text{if $p_D(2015) \geq 0.9$}.
      \end{cases}
      \\
      p_T^*(t) &=
      \begin{cases}
        F\big(t, 2015, p_T(2015), 2020, 0.9\big)
        & \text{if $p_T(2015) < 0.9$},
        \\
        p_T(2015) & \text{if $p_T(2015) \geq 0.9$}.
      \end{cases}
      \\
      p_V^*(t) &=
      \begin{cases}
        F\big(t, 2015, p_V(2015), 2020, 0.9\big)
        & \text{if $p_V(2015) < 0.9$},
        \\
        p_V(2015) & \text{if $p_V(2015) \geq 0.9$}.
      \end{cases}
    \end{split}
  \end{equation}
  where
  \begin{equation}
    \label{F}
    F(t, t_0, x_0, t_1, x_1) =
    \begin{cases}
      x_0 & \text{if $t < t_0$},
      \\
      x_0 + (x_1 - x_0) \frac{t - t_0}{t_1 - t_0} &
      \text{if $t_0 \leq t < t_1$},
      \\
      x_1 & \text{if $t \geq t_1$},
    \end{cases}
  \end{equation}
  is constant at $x_0$ for $t < t_0$, linearly connects $x_0$ at
  $t_0$ to $x_1$ at $t_1$ for $t_0 \leq t < t_1$, and is constant at
  $x_1$ for $t \geq t_1$.

\item[UNAIDS 95--95--95] The UNAIDS 95--95--95 target is to achieve
  90--90--90 by 2020 and then for $p_D$, $p_T$, and $p_V$ to all be
  raised to $95\%$ by 2030.  We implemented this target as for
  90--90--90 from 2015 to 2020, and then linear increases from the
  2020 levels up to $95\%$ in 2030, again with the exception that if
  an initial level is above $95\%$ then it remains constant from 2015
  to 2035.  That is,
  \begin{equation}
    \label{unaids95_targets}
    \begin{split}
      p_D^*(t) &=
      \begin{cases}
        G\big(t, 2015, p_D(2015), 2020, 0.9, 2030, 0.95\big)
        & \text{if $p_D(2015) < 0.9$},
        \\
        F\big(t, 2020, p_D(2015), 2030, 0.95\big)
        & \text{if $0.9 \leq p_D(2015) < 0.95$},
        \\
        p_D(2015) & \text{if $p_D(2015) \geq 0.95$},
      \end{cases}
      \\
      p_T^*(t) &=
      \begin{cases}
        G\big(t, 2015, p_T(2015), 2020, 0.9, 2030, 0.95\big)
        & \text{if $p_T(2015) < 0.9$},
        \\
        F\big(t, 2020, p_T(2015), 2030, 0.95\big)
        & \text{if $0.9 \leq p_T(2015) < 0.95$},
        \\
        p_T(2015) & \text{if $p_T(2015) \geq 0.95$},
      \end{cases}
      \\
      p_V^*(t) &=
      \begin{cases}
        G\big(t, 2015, p_V(2015), 2020, 0.9, 2030, 0.95\big)
        & \text{if $p_V(2015) < 0.9$},
        \\
        F\big(t, 2020, p_V(2015), 2030, 0.95\big)
        & \text{if $0.9 \leq p_V(2015) < 0.95$},
        \\
        p_V(2015) & \text{if $p_V(2015) \geq 0.95$},
      \end{cases}
    \end{split}
  \end{equation}
  where
  \begin{equation}
    G(t, t_0, x_0, t_1, x_1, t_2, x_2) =
    \begin{cases}
      x_0 & \text{if $t < t_0$},
      \\
      x_0 + (x_1 - x_0) \frac{t - t_0}{t_1 - t_0} &
      \text{if $t_0 \leq t < t_1$},
      \\
      x_1 + (x_2 - x_1) \frac{t - t_1}{t_2 - t_1} &
      \text{if $t_1 \leq t < t_2$},
      \\
      x_2 & \text{if $t \geq t_2$},
    \end{cases}
  \end{equation}
  is constant at $x_0$ for $t < t_0$, linearly connects $x_0$ at $t_0$
  to $x_1$ at $t_1$ for $t_0 \leq t < t_1$, linearly connects $x_1$ at
  $t_1$ to $x_2$ at $t_2$ for $t_1 \leq t < t_2$, and is constant at
  $x_2$ for $t \geq t_2$, and $F$ is as in \eqref{F}.

\end{description}

For vaccination, we assume that vaccination rollout begins in year
$t_V = 2020$ (or $t_V = 2025$ in vaccine scenario analysis,
\autoref{uncertainty}), that rollout linearly increases the proportion
of non-infected population covered from 0\% by $r_V = 25\%$ per year
(or $r_V = 10\%$ per year) up to $c_V = 70\%$ (or $c_V = 50\%$ or
$c_V = 90\%$) and staying constant thereafter.  Thus, the vaccination
target is
\begin{equation}
  \label{vaccination_target}
  p_R^*(t) =
  \begin{cases}
    0 & \text{if $t < t_V$},
    \\
    r_V (t - t_V) & \text{if $t_V \leq t < T_V$},
    \\
    c_V & \text{if $t \geq T_V$},
  \end{cases}
\end{equation}
where $T_V = t_V + c_V / r_V$ is the time when rollout reaches
the full coverage $c_V$.

\section{Numerical solution}

For accuracy and speed in numerically solving differential equations
\eqref{model_eqns}, the variables $S$, $U$, $D$, $T$, $V$, and
$W$---i.e., those that are not initially $0$---were log transformed
and the differential equations were written in terms of
the new log variables:
\begin{equation}
  \label{model_eqns_log}
  \begin{split}
    \frac{\md \log S}{\md t} &= \kappa N \me^{- \log S} - \lambda - \zeta - \mu,
    \\
    \frac{\md R}{\md t} & = \zeta \me^{\log S} - (1 - \rho) \lambda R - \mu R,
    \\
    \frac{\md A}{\md t} &= \lambda \me^{\log S} + (1 - \rho) \lambda R
    - \delta A - \mu A,
    \\
    \frac{\md \log U}{\md t} &= \delta A \me^{- \log U} - \alpha - \mu - \sigma,
    \\
    \frac{\md \log D}{\md t} &=  \alpha \me^{\log U - \log D}
    + \psi \me^{\log T - \log D} + \psi \me^{\log V - \log D} - \phi - \mu - \sigma,
    \\
    \frac{\md \log T}{\md t} &= \phi \me^{\log D - \log T} - \psi - \gamma - \mu
    - \sigma,
    \\
    \frac{\md \log V}{\md t} &= \gamma \me^{\log T - \log V} - \psi
    - \mu  - \theta,
    \\
    \frac{\md \log W}{\md t} &= \sigma \me^{\log U - \log W}
    + \sigma \me^{\log D - \log W} + \sigma \me^{\log T - \log W}
    + \theta \me^{\log V - \log W} - \nu,
    \\
    \frac{\md Y}{\md t} &= \lambda \left[\me^{\log S} + (1 - \rho) R\right],
    \\
    \frac{\md Z}{\md t} &= \nu \me^{\log W}.
  \end{split}
\end{equation}

The model simulation and analysis tools, written in Python, are
publicly available at
\url{https://github.com/janmedlock/HIV-95-vaccine/}
\cite{medlock2016-git}.  The differential equations were solved using
\texttt{scipy.integrate.odeint()} \cite{scipy}.  See the source code
for full details.

@@

\comment{Vaccine scenarios.}
